Real-time resolving assembly graph by ONT long reads data

### 2019-11-30 npgraph.tex
create npgraph paper

\usepackage{pdfpages} 
\usepackage{mathtools}
\usepackage{play}
\usepackage{makeidx}
\usepackage{xcolor,colortbl}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{pdflscape}

@@

\title{Real-time resolving assembly graph by ONT long reads data}
\shorttitle{Real-time Scaffolding with Assembly graph}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Introduction}
% \begin{itemize}
% \item The importance of having a real-time assembler for bacterial and small genomes: e.g. npscarf 
% \item The difficulty to resolve assembly graph in real-time.
% \item Available hybrid assembly methods using long reads: unicycler, ...
% \item an online algorithm to resolve assembly graph by long reads is needed, ...
% \item Requirements
%     \begin{itemize}
%     \item Support streaming: efficient with high yield (armortized analysis)
%     \item Statistically robust: online modification of the dynamic graph, back tracking edge deletion...
%     \item Standardisation : output GFA in real-time
%     \end{itemize}
% \end{itemize}
Streaming assembly methods had been proven to be useful in saving time and resources compared to the conventional batch algorithms with examples included \EG{} $\mathtt{Faucet}$~\cite{Rozov2017faucet} and \npscarf{}~\cite{Cao2017scaffolding}. The first method allows the assembly graph to be constructed incrementally as long as reads are retrieved and processed. This practice is helpful dealing with huge short-read data set because it can significantly reduce the local storage for the reads, as well as save time for a De Bruijn graph (DBG) construction while waiting for the data being retrieved.
\npscarf{}, on the other hand, is a hybrid assembler working on a pre-assembly set of short-read assembly contigs. It functions by scaffolding the contigs using nanopore sequencing which is well-known by the real-time property. The completion of genome assembly in parallel with the sequencing run provides explicit benefits in term of resource control and turn-around time for analysis~\cite{Cao2017scaffolding}.

We use different hybrid data sets to study the \npgraph{}`s performance and compare with other available tools with similar functionality, such as \npscarf{}, \unicycler{} and \hspades{}. Even though all data were from microbial genomes, the complexity of the input graphs range from few hundred to ten thousands of components, synthesized \emph{in silico} or sequenced from \emph{in vitro} samples.
%Particularly, for the first time, we employ a pipeline in an attempt to complete metagenomics assembly in real-time. 
%We aim to develop a simple approach on top of \npgraph{} to adapt with metagenomics while minimizing additional processing steps. In fact, only one more step for population binning is needed to make \npgraph{} functions on genomes of a mock community.
%Despite a rather straightforward methodology, the pipeline has shown consecutively enhanced assembly from the fragmented short-read DBG graph.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
\subsection*{\npgraph{} -- a tool to resolve assembly graph in real-time}
\npgraph{}'s input consists of Illumina assembly graph resulted from running assembler, e.g. \spades{}~\cite{BankevichNA2012}, $\mathtt{Velvet}$ \cite{Zerbino2008}, $\mathtt{AbySS}$ \cite{Simpson2009} on Illumina short reads, together with long reads from third generation sequencing technology (Oxford Nanopore Technology, Pacbio).
The long reads will be aligned with the contigs in the assembly graph to indicate longer paths that should be traversed. These local paths, given sufficient data, are expected to untangle the complicated graph and guide to the global Eulerian paths (or cycles if possible) that represent the entire genomic sequences. 
\npgraph{} can be invoked and fully function from the command-line interface. In addition, in order to aid the visualization of the assembly process, a GUI has been developed as well.

@@

The GUI includes the dashboard for control the settings of the program and another pop-up window for a simple visualization of the assembly graph in real-time (Figure~\ref{figure:npgraph_gui}).
In this interface, the assembly graph loading stage is separated from the actual assembly process so that users can check for the graph quality first before carry out any further tasks. The box numbered \textbf{1} on Figure~\ref{figure:npgraph_gui} is designed for this task.
Only after an assembly graph is loaded successfully, users can move to box \textbf{2} to specify the nanopore input data.
Settings for an aligner (\bwa{} or \minimap{}) in box \textbf{3} is required if the input is the raw sequences in FASTA/FASTQ format. Another option is to run the alignment independently and provide SAM/BAM input for the next stage of bridging and assembly. This stage is controlled by buttons in box \textbf{4}: the START button ignites the process while the STOP button can prematurely terminate it and output the assembly result till that moment. The plots from the right panel (\textbf{5}) depicts real-time statistics of the assembly contigs inferred from the graph.
From the second window (\textbf{6}), the colored vertices imply unique contigs while the white ones involve either unspecified or repetitive elements. The number of different colors (other than white) indicates the amount of abundant groups being detected as population bins (\EG{} chromosome versus different plasmids, or different bins in metagenomics).

A proper combination of command line and GUI can provide an useful streaming pipeline that copes well with MinION output data. The practice is similar to the previous developed pipelines~\cite{CaoGC2016,Cao2017scaffolding,Nguyen2017barcode} that allow the analysis to take place abreast to a nanopore sequencing run.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Conclusion}
%npgraph: how important hybrid assembly
Due to the limits of current sequencing technology, application of hybrid methods should remain a common practice in whole genome assembly for the near future.
On the other hand, the ONT platforms are evolving quickly with significant improvement in terms of data accuracy and yield, however, the sequencing cost per base is still high. Beside, the real-time property of this technology has not been sufficiently exploited to match its potential benefits.
\npscarf{} had been introduced initially to address these issues, however, the accuracy of the assembly output was affected by its greedy alignment-based scaffolding approach.
Here we present \npgraph{}, a streaming hybrid assembly method working on the DBG assembly graph that is able to finish short-read assembly in real-time while minimizing the errors and mis-assemblies drastically.

>>>>

\usepackage{pdfpages} 
\usepackage{mathtools}
%\usepackage{play}
\usepackage{makeidx}
\usepackage{xcolor,colortbl}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{pdflscape}

@@

\title{
Real-time resolving assembly graph by ONT long reads data}
\shorttitle{Real-time Scaffolding with Assembly graph}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Introduction}

%Highlight the importance of real-time/hybrid assembly

%Gaps: Several tool apptemt to use long reads to

% \begin{itemize}
% \item The importance of having a real-time assembler for bacterial and small genomes: e.g. npscarf 
% \item The difficulty to resolve assembly graph in real-time.
% \item Available hybrid assembly methods using long reads: unicycler, ...
% \item an online algorithm to resolve assembly graph by long reads is needed, ...
% \item Requirements
%     \begin{itemize}
%     \item Support streaming: efficient with high yield (armortized analysis)
%     \item Statistically robust: online modification of the dynamic graph, back tracking edge deletion...
%     \item Standardisation : output GFA in real-time
%     \end{itemize}

There is rich  connectivity information in the assembly graph resulted from short read assembly. However, existing tools
do not make use of this. Here, we present npGraph, a novel algorithm to resolve the assembly graph in real-time long read 
sequyencing. npGraph maintain the assembly grahp and as new long reads coming in, it untangle the knots thereby simplify
the assembly graph. Because of this, npGraph has better estimation of multiplcity of repeat contigs and hence resulting in
fewer misassembly (Not to mention in real-time). In addition, we develop a visualisation tool for practitioners to monitor
the progress of assembling process.

.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
\subsection*{Resolving assembly graph in real-time}

\npgraph{} makes use of a assembly graph resulted from assembling short reads using a de Bruijin
graph method such as \spades{}~\cite{BankevichNA2012}, $\mathtt{Velvet}$~\cite{Zerbino2008}
and $mathtt{AbySS}$~\cite{Simpson2009}. The assembly graph consists of a list of contigs, and 
possible connections among these contigs. In building the assembly graph, the de Bruijin graph assembler attemtps to extend each contig as long as possible, until there are more than one possible
ways of extending due to the repeatitive sequences beyond the informaion contained in short reads. Hence cach contig has multiple possible connections with others, creating knots in the assembly graph. \npgraph{} then uses the connectivity information from long reads to untangle these knots in real-time. With sufficient data, when all the knots are removed, the assemply graph is simplified 
to a path which represents the complete assembly.

\npgraph{} aligns long reads to the contigs in the assembly graph. When a long read is aligned
to multiple contigs, \npgraph{} constructs candidate paths that are supported by the read. This 
stratery allows \npgraph{} to progressively update the likelihood of the paths going through a knot.
When sufficient data are obtained, the best path is confidently identified and hence the knot is untangled.
%TODO: two more sentences describing the bridging
%TODO: Because of the converting long reads to contig alignment to mapping to path, npgraph is faster than unicycler, and can be done in real-time

% 
% These local paths, given sufficient data, are expected to untangle the complicated graph and guide to the global Eulerian paths (or cycles if possible) that represent the entire genomic sequences. 
% \npgraph{} can be invoked and fully function from the command-line interface. In addition, in order to aid the visualization of the assembly process, a GUI has been developed as well.

@@

We also develop a Graphical User Interface (GUI) for npgraph. The GUI includes the dashboard for controling the settings of the program and a window for visualization of the assembly graph 
in real-time (Figure~\ref{figure:npgraph_gui}). 
In this interface, the assembly graph loading stage is separated from the actual assembly process so that users can check for the graph quality first before carry out any further tasks.
A proper combination of command line and GUI can provide an useful streaming pipeline that copes well with MinION output data. The practice is to support the real-time monitoring of
the results from real-time sequencing~\cite{CaoGC2016,Cao2017scaffolding,Nguyen2017barcode} that allow the analysis to take place abreast to a nanopore sequencing run.

% TODO: Extract information from here to caption
% The box numbered \textbf{1} on Figure~\ref{figure:npgraph_gui} is designed for this task.
% Only after an assembly graph is loaded successfully, users can move to box \textbf{2} to specify the nanopore input data.
% Settings for an aligner (\bwa{} or \minimap{}) in box \textbf{3} is required if the input is the raw sequences in FASTA/FASTQ format. Another option is to run the alignment independently and provide SAM/BAM input for the next stage of bridging and assembly. This stage is controlled by buttons in box \textbf{4}: the START button ignites the process while the STOP button can prematurely terminate it and output the assembly result till that moment. The plots from the right panel (\textbf{5}) depicts real-time statistics of the assembly contigs inferred from the graph.
% From the second window (\textbf{6}), the colored vertices imply unique contigs while the white ones involve either unspecified or repetitive elements. The number of different colors (other than white) indicates the amount of abundant groups being detected as population bins (\EG{} chromosome versus different plasmids, or different bins in metagenomics).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}

% \end{itemize}
Streaming assembly methods had been proven to be useful in saving time and resources compared to the conventional batch algorithms with examples included \EG{} $\mathtt{Faucet}$~\cite{Rozov2017faucet} and \npscarf{}~\cite{Cao2017scaffolding}. The first method allows the assembly graph to be constructed incrementally as long as reads are retrieved and processed. This practice is helpful dealing with huge short-read data set because it can significantly reduce the local storage for the reads, as well as save time for a De Bruijn graph (DBG) construction while waiting for the data being retrieved.
\npscarf{}, on the other hand, is a hybrid assembler working on a pre-assembly set of short-read assembly contigs. It functions by scaffolding the contigs using nanopore sequencing which is well-known by the real-time property. The completion of genome assembly in parallel with the sequencing run provides explicit benefits in term of resource control and turn-around time for analysis~\cite{Cao2017scaffolding}.

We use different hybrid data sets to study the \npgraph{}`s performance and compare with other available tools with similar functionality, such as \npscarf{}, \unicycler{} and \hspades{}. Even though all data were from microbial genomes, the complexity of the input graphs range from few hundred to ten thousands of components, synthesized \emph{in silico} or sequenced from \emph{in vitro} samples.
%Particularly, for the first time, we employ a pipeline in an attempt to complete metagenomics assembly in real-time. 
%We aim to develop a simple approach on top of \npgraph{} to adapt with metagenomics while minimizing additional processing steps. In fact, only one more step for population binning is needed to make \npgraph{} functions on genomes of a mock community.
%Despite a rather straightforward methodology, the pipeline has shown consecutively enhanced assembly from the fragmented short-read DBG graph

\section*{Conclusion}
%npgraph: how important hybrid assembly
Due to the limits of current sequencing technology, application of hybrid methods should remain a common practice in whole genome assembly for the near future.
On the other hand, the ONT platforms are evolving quickly with significant improvement in terms of data accuracy and yield, however, the sequencing cost per base is still high. Beside, the real-time property of this technology has not been sufficiently exploited to match its potential benefits.
\npscarf{} had been introduced initially to address these issues, however, the accuracy of the assembly output was affected by its greedy alignment-based scaffolding approach.
Here we present \npgraph{}, a streaming hybrid assembly method working on the DBG assembly graph that is able to finish short-read assembly in real-time while minimizing the errors and mis-assemblies drastically.
